---
title: "Logicals"
author: "Juliet"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

- if you write include = FALSE in the yaml, the whole chunk does not show up
- if you include echo = FALSE, none of the code will show up in the knitted HTML but the output will
- eval = FALSE tells R to not run that chuck


## Taking a look at star wars

```{r, echo = FALSE}

#if you write include = FALSE in the yaml, the whole chunk does not show up
#if you include echo = FALSE, none of the code will show up in the knitted HTML but the output will
summary(starwars)
head(starwars)
tail(starwars)
dim(starwars)
```


```{r, eval = FALSE}
ggplot(data = starwars, aes(x = height, y = mass)) +
  geom_point(aes(color = species))
#use aes() here because species is based on a variable, if you are just requesting the points to be a color you dont use aes()
#warning = FALSE in the yaml means that the warning R gives you does not show up in the knitted document
```

# calculations on a single column

```{r}
sw_height_mean_try1 <- mean(starwars$height)
#returns NA because there are missing values and the mean function does not automatically remove NA values, this is a goof thing, makes you take a second look at your data
sw_height_mean_try2 <- mean(starwars$height, na.rm = TRUE)
sw_height_mean_try2

#summary(starwars$height)

#using similar naming system, find:
#variance of character height (var())
#std of height (sd))

sw_charac_height_var <- var(starwars$height, na.rm = TRUE)
sw_charac_height_var

sw_charac_height_sd <- sd(starwars$height, na.rm = TRUE)
sw_charac_height_sd
```

## Report stats in R rather than typing them put manually, which is prone to mistakes

Mean character height is `r round(sw_height_mean_try2, 1)` cm, with a standard deviation of `r roudn(sw_charac_height_sd, 1)`

## Logical Expressions
```{r}
5 < 10

"cat" > "armadillo"
#returns true because R works in alphabetical order with a being a lower value than c

"cats" > "cat"
#returns true because cats has a higher value since R uses the "assy" code to assign numebrs to letters

"dog" == "dog"


```

```{r}
10 <= 11

5 >= 4

3 <= 1

10 == 11
```


```{r}
tarantulas <- c(1, 3, 7, 10, 15)
tarantulas > 5


```

# logical operators

- `!`
- `&` (and statement)
- `|` (or statement)

```{r}
3 < 1 | 8 > 2

6 < 10 & 5 < 1

7 != 11
#7 is not equal to 11

4 != 4 | 2 < 5 | 6 > 10

0 == FALSE
1 == TRUE
#if you have a binary variable stored as 0's and 1's, R will automatically treat them as false or true respectively 
```




